### 2016-06-29 paper/abstract.tex
added in fixes yadda ydadda
abstract

The language of lenses expresses programs which can convert between source and
target regular languages.
We tackle the task of synthesizing these programs through the regular expressions
of their types.
This is difficult because we cannot use basic type directed synthesis techniques,
as there are many equivalences in the types of these regular expressions.
We do this by developing a language for partially normalized regular expressions,
and searching through those regular expressions for two that have a lens between
them which can be synthesized in a type-directed way.
We develop strategies for efficiently enacting this search and synthesis,
and evaluate our strategies against a suite of microbenchmarks and linux
configuration files.

>>>>

Tools for reliably converting back and forth between different data
formats are increasingly important in a world full of big data.
One class of such tools are \emph{lenses}, programs that simultaneously
represent one function to convert data from type $A$ to type $B$ and a
second function to convert data from type $B$ back into data of type
$A$.  While such programs often come with important invertability properties,
they can be quite difficult to write in practice 
as a programmer must think about data conversions in two directions as once.
To overcome this difficulty, we develop algorithms for synthesizing such lenses 
automatically from their types and a small number of examples.
Unlike past work in type-directed synthesis, synthesis of lenses is
complicated by their rich type structure, which is based on regular
expressions.  More specifically, rather than simply having to solve
the problem of finding a representative term to map back and forth
between types $A$ and $B$, it is first necessary to find the right
representative types $A'$ and $B'$ amongst many equivalent regular expressions.
We show how to manage this search by defining a new language of
lenses governed by semi-canonical regular expression types and prove it
sound with respect to a standard language of bijective lenses.  We then show
how to search the reduced space for well-typed lenses that are consistent
with the given examples.  We demonstrate the effectiveness of our algorithms
by experimenting on benchmark suite containing microbenchmarks, FlashFill
examples, and linux configuration file transformations.

%% The language of regular lenses expresses programs which can convert between
%% source and target regular languages.
%% We tackle the task of synthesizing these programs through the regular expressions
%% of their types.
%% This is difficult because we cannot use basic type directed synthesis techniques,
%% as there are many equivalences in the types of these regular expressions.
%% We do this by developing a language for partially normalized regular expressions,
%% and searching through those regular expressions for two that have a lens between
%% them which can be synthesized in a type-directed way.
%% We develop strategies for efficiently enacting this search and synthesis,
%% and evaluate our strategies against a suite of microbenchmarks and linux
%% configuration files.
